Attributions for Poverty in Post-Socialist Countries

### 2014-02-21 article2013.Rmd
5.3. muutoksia alkuun, jatkuu vielä tänään

As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables. Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.

@@

Tables 3, 4 and 5 present figures for multilevel logistic regressions. Calculating R squared or residual standard deviation makes no sense in the case of logistic regression or other binomial models. Instead, we use *deviance* as our primary statistical summary of logistic model fit. Deviance can be treated as similar to residual standard deviation for generalized linear models. Basically, deviance is a measure of error, and lower deviance implicates better fit to data. When a random predictor is added to model, we expect deviance to decrease, on average, by 1. When predictor is more informative deviance should decrease more than 1. (Gelman and Hill, 2007). As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.

Between-country variances are on the first rows in random effects, meaning the variance in dependent variable that is explained by country differences. Variance in empty models is .29 in social blame and .12 in individual blame indicating that there are differences between countries, and that differences are greater in respect to support for social blame type of explanation. Inclusion of individual level predictors decreases the variance in both models, but only moderately.

Looking at the deviance for the empty models both in tables 3 and 4, we can see that there is "less" to be explained for in the model for individual blame than for social blame. When adding the individual level predictors to the models, the deviance decreases by 8574 for social blame and 6662 for individual blame. In both model *SB individual* and *IB individual* we have four statistically significant predictors. *Perceived low income level* has the highest coefficient of our binary individual level predictors for explaining social blame type of explanation. Coefficient of .36 can be translated so that if respondent perceives belonging to low income group he or she is 8.5 % more likely to blame society for poverty than a non-low income individual. Expected worsening in household income and financial crises experince predict social blame almost in similar manner. *Low education* and *worsened household income* do not contribute in explaining neither of the outcomes.  Being greatly affected by financial crisis makes one 11 % less likely to support for individual blame type of explanation.

>>>>

Descriptive analyses presented above would posit that there are rather clear, though to some extent unexpected, associations between country-level contextual characteristics and the perceptions of the causes of poverty. In order to explore the effects of both contextual and individual-level determinants in a more detailed statistical manner, the analytsis continues with multilevel logistic regressions (Tables 3, 4 and 5).

Between-country variances in Tables 3 and 5 are indicators for the variance in dependent variable that is explained by country differences. Variance in empty models is .29 in social blame and .12 in individual blame indicating that there are differences between countries, and that differences are greater in respect to support for social blame type of explanation. Inclusion of individual-level predictors decreases the variance in both models, but only moderately.

In addition, while calculating R squared or residual standard deviation makes no sense in the case of binomial models, our primary statistical summary of logistic model fit is *deviance*. Deviance can be treated as similar to residual standard deviation for generalized linear models. Basically, deviance is a measure of error, and lower deviance implicates better fit to data. When a random predictor is added to model, we expect deviance to decrease, on average, by 1. When predictor is more informative deviance should decrease more than 1. (Gelman and Hill, 2007). As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate. The deviance for the empty models both in Tables 3 and 4 indicates that there is "less" to be explained for in the model for individual blame than for social blame. When adding the individual-level predictors to the models, the deviance decreases by 8574 for social blame and 6662 for individual blame.

@@

 In both model *SB individual* and *IB individual* we have four statistically significant predictors. *Perceived low income level* has the highest coefficient of our binary individual level predictors for explaining social blame type of explanation. Coefficient of .36 can be translated so that if respondent perceives belonging to low income group he or she is 8.5 % more likely to blame society for poverty than a non-low income individual. Expected worsening in household income and financial crises experince predict social blame almost in similar manner. *Low education* and *worsened household income* do not contribute in explaining neither of the outcomes.  Being greatly affected by financial crisis makes one 11 % less likely to support for individual blame type of explanation.
